Emulsions Stabilized by Magnetic Ellipsoidal Particles: A Lattice Boltzmann Study

### 2025-04-06 chapters/paper2.tex
Changes 1. Rewrote all of aim 2 except conclusion

\section{Results}\label{sec:results_p2}
\subsection{Hysteresis curve}\label{section:hysteresis_curve}

Bijels have proposed applications as soft matter based material templates, drug delivery vectors and separations systems. The materials 
performance in each application is dependent upon the microstructure of the bijel template. Many of the applications listed above benefit 
from stimuli response, allowing for in-situ microstructure tuning to offer reversible modifications of material permeability. In this section, we analyze
whether stimuli responsive bijels provide access to reversible structural modifications.

We begin our investigation by characterizing the microstructure changes observed on bijels stabilized with prolate ellipsoids with a 
particle volume fraction of $\phi_p = 0.1$. We only focus on one system here as the mechanism for stimuli response shown in Aim 1 is identical
for both systems. Therefore We expect that the qualitative trends between prolate and oblate particles will be matched. We select prolate particles
as they are more readily synthesizable in experiments.

We apply a constant magnetic field of $\bar{B}_z = 0.2$ onto bijels that were simulated 
with no magnetic field. Once the microstructure has reached its new state, we ramp up the magnetic field to $\bar{B}_z = 0.5$ and 
finally $\bar{B}_z = 1$. We then apply the opposite procedure to reduce the applied field strength, resulting in a magnetic field 
ramp of $\bar{B}_z: 0 \rightarrow 0.2 \rightarrow 0.5 \rightarrow 1 \rightarrow 0.5 \rightarrow 0.2 \rightarrow 0$. We characterize the 
microstructure changes observed through calculating the domain size, defined as the first moment of the spherically averaged structure factor. 
We calculate the hysteresis 
curve by selecting the domain size at the final timestep of each magnetic field change. We then plot these domain sizes against the magnetic field 
applied to obtain them and split them in the descending and ascending magnetic field directions to demonstrate the differences observed. We plot 
these results in Figure \ref{fig:hysteresis_curve}

@@

Figure \ref{fig:hysteresis_curve} demonstrates the hysteresis response of the bijel microstructure, characterized by a difference in the average domain size
when increasing and decreasing the magnetic field. The increase in the average domain size when applying the magnetic field is due to domain coarsening caused by 
reorientation of particles to the magnetic field strength, reducing the interfacial coverage. When decreasing the magnetic field strength, 
the domain size does not return to the original value meaning that that the microstructure of the bijel is unaffected by 
the reduction and eventual removal of the applied magnetic field.

This behavior is also observed in past work investigating the shape deformation of particle stabilized emulsions under 
electric fields. \cite{cui_stabilizing_2013} Cui et al. demonstrated that the particles unjammed and rejammed into a new kinetically arrested state resulting in
the droplet being elongated in the direction of the applied field. This suggests that the particles at the interface of magnetically responsive bijels are in
in a kinetically trapped configuration. Cui et al. used spherical particles in their emulsion response experiments while we use ellipsoidal particles. Additionally,
Figure \ref{fig:hysteresis_curve} shows that the response of the bijel is dependent upon the field strength applied.
In the following sections, we investigate the impact of magnetic field strength and the initial ordering of the particle monolayer on the observed stimuli response.

\subsection{Field strength dependence on domain size}
\label{section:field-strength-dependence-on-domain-size}

From the results in Figure \ref{fig:hysteresis_curve}, we show that the domain size change is dependent upon the applied magnetic field. As porous materials templates,
the microstructure of the material is of interest as it controls the performance of the material. Additionally, in many of the proposed stimuli responsive applications
for bijels, the timescales of response is important to characterize and would also provide a mechanistic understanding of the reasons why the microstructure changes 
occur. In this section, we analyze the structural response of bijels by assessing the response of bijels stabilized by particles with volume fraction of 
$\phi_p = 0.1$ simulated under no applied field to an applied field in the range $\bar{B} = 0, 0.2, 0.5, 1$. We begin our investigation by visualizing the structure before 
and after the application of magnetic fields.

\begin{figure} 
\centering 
\includegraphics[scale=0.4]{../figures/results/paper2/microstructure_viz-field_on.png} 
\caption{Visualizations of bijels stabilized by oblate and prolate particles simulated under no fields at $t = 0$ (left columns) and $t = 10^5$ (right columns). The first 
         and third row detail the microstructure evolution for oblate and prolate particle stabilized bijels with no applied field while the second and fourth rows show 
         bijels stabilized by oblate and prolate particles respectively with a field strength of $\bar{B} = 1$ applied. Particle reorientation to the direction of the 
         field can be seen upon application of the magnetic field, resulting in microstructure changes driven through domain coarsening until the particles jam. We see 
         similar behavior in bijels stabilized by oblate particles upon application of the magnetic field.}
\label{fig:microstructure_viz-field_on} 
\end{figure}

In Figure \ref{fig:microstructure_viz-field_on} we show representative snapshots of the bijel at the initial and final timestep with $\bar{B}_z = 0$ and $\bar{B}_z = 1$. 
Upon application of the field, particles align to the direction of the field. As the particles orient to the field, they disrupt the packing of particles at
the interface. This causes domain coarsening as the particles rearrange into new positions before rejamming in place once the interfacial
area and particle cross sectional area match. Domain coarsening observed in the figure can be explained through the ordering of particles at the interface, causing a 
reduction in the available area that the particles have for stabilization. To characterize the length scale of the microstructure upon application of the magnetic field,
we calculate the domain size. To correlate the microstructure evolution to the ordering of the particle monolayer to the magnetic field, we overlay the time evolution of the
nematic order parameter on the domain size. \cite{veerman_phase_1992}

\begin{figure} 
\centering 
\includegraphics[scale=0.5]{../figures/results/paper2/domain_size-field_on.png} 
\caption{Plotting the spherically averaged domain size normalized with $R_p$($L/R$) of the particle at different magnetic field strengths, $\bar{B}$, and the nematic order 
         parameter $S$ in lines. We quantify the domain size and ordering of the particles, and show that the change in both parameters are correlated with the strength of 
         the applied magnetic field.} 
\label{fig:domain_size-field_on} 
\end{figure}

We normalized the domain size results by the volume equivalent sphere radius, $R = 7.9$ in Figure \ref{fig:domain_size-field_on} The results indicate that upon 
applying a magnetic field to a bijel initially formed without one, the final average domain size increases by up to 5.2\% and 2.5\% for oblate and 
prolate particle stabilized bijels respectively. The domain coarsening is initially slow, before accelerating rapidly and finally reaching a maximum where it plateaus. 
Additionally, in the absence of an external field, domain coarsening occurs for both particle morphologies. However, the results
for oblate particles with a magnetic field is attributed to coalescence of domains rather than magnetic field driven ordering of the particles.

In the same figure, the time evolution of the nematic order parameter $S$ is also plotted to illustrate changes in particle alignment.
Initially, at $\bar{B}_z = 0$, a gradual increase in $S$ is observed, attributed to steric interactions driving particle reordering.
This effect contributes to the domain coarsening seen in the absence of an applied field, consistent with the findings of Günther et
al. \cite{gunther_timescales_2014}. In the case of oblate particle stabilized bijels without an external field, domain
coalescence is primarily driven by steric interactions rather than field-induced rearrangement. Furthermore, the rate and extent of
particle alignment to the field exhibit a dependence on field strength, as evidenced by the earlier onset of nematic ordering with increasing
$\bar{B}$. A time delay between the evolution of the nematic order parameter and domain size, particularly when comparing results for
$\bar{B}_z = 0.5, 1$ against those for $\bar{B}_z = 0.2$ are seen. This suggests that nematic ordering alone may not fully account for the
observed changes in domain size.

Previous studies have demonstrated that microstructural properties such as the length scale and tortuosity, play an important role in the
performance of porous materials across various applications, including battery electrodes and tissue engineering. \cite{zhang_promoting_2019}
\cite{ebner_tortuosity_2014} \cite{hsieh_architected_2021} Specifically, lower tortuosity and more ordered ion-conducting domains have been shown
to enhance charge/discharge rates in battery systems, while controlled tortuosity in artificial bone grafts facilitates improved cell growth
and nutrient transport.

Precise characterization of the anisotropic microstructure is essential for the tailored design of application-specific porous materials. To
quantify the characteristic length scales in each Cartesian direction, we compute the second-moment-derived structure factor. \cite{jansen_bijels_2011} 
We make a further refinement simplifying the domain sizes to directions parallel and orthogonal to the magnetic field. These result in the length scales and
their associated calculation, $L_{\parallel} = L_{z}$ and $L_{\perp} = \frac{L_x + L_y}{2}$. The same simplification
with the directional tortuosities is also performed.

Tortuosity has been defined using the geometric, diffusional and hydraulic tortuosities. The tortuosity calculated here is
the diffusional tortuosity, $\tau = \frac{D}{D_{eff}}$ which characterizes the ratio between the free diffusion of a tracer and the
effective diffusion of a tracer in the material. We plot the directional second moment derived length scales and tortuosity in
Figure \ref{fig:domain_size_aniso-field_on}.

@@

From Figure \ref{fig:domain_size_aniso-field_on}, domain size anisotropy is observed upon application of the field for both
particle morphologies. Some domain anisotropy with no applied field is observed, which increases upon application of the magnetic field.
The directional domain sizes, $L_{\perp}$ and $L_{\parallel}$ increase by 73\% and 7\%\ respectively for oblate particles and
10\% and 44\% respectively for oblate particles. The axes that coarsens the most is consistent with past work, explained through the
directions particles reorient in response to the applied magnetic field. As particles unjam and rejam at the interface due to reorientating 
to the magnetic field, they adopt direction specific surface coverages which manifest as anistropic domains upon jamming. In Figure 
\ref{fig:domain_size_aniso-field_on}, we demonstrate that the application of the magnetic field changes the domain size.

The tortuosity of the bijel is initially around $\tau \approx 1.5$, consistent with simulations of the diffusive tortuosities of gyroid structures using 
Lattice Boltzmann simulations. \cite{luo_macroscopic_2020} Upon application of the field, $\tau_{\perp}$ slightly decreases while $\tau_{\parallel}$
increases. In Chapter \ref{chapter:aim1}, we characterized that increasing the domain size decreases the tortuosity. \cite{karthikeyan_formation_2024} This
observation is in line with the domain size changes observed in Figure \ref{fig:domain_size_aniso-field_on}. The results show that the
tortuosity of the bijel is noticeably affected by the application of magnetic fields. This data suggests that magnetic fields applied onto
bijels stabilized with ellipsoidal particles undergo domain anisotropy upon application of magnetic fields. However the particle morphology,
specifically its axis of symmetry, affect the direction that the domain anisotropy is characterized.

Next, we investigate the time gap between the average domain size increase and the nematic order parameter. This result suggests that the dynamics 
of the system at higher magnetic field strengths are dependent upon other factors that vary upon application of the magnetic field. In this system, these factors
would arise from capillary forces with the interface or steric interactions with other particles at the interface, slowing down the dynamics of the system.
To understand this phenomena, we focus on the dynamics of the particle monolayer. First, we visualize the particle monolayer of the bijels stabilized with prolate 
ellipsoids at the initial and final configuration before and after application of a magnetic field.

\begin{figure} 
\centering 
\includegraphics[scale=0.2]{../figures/results/paper2/particle_viz-field_on.png} 
\caption{Visualizations of bijels stabilized with oblate and prolate particles simulated at $t = 0$ (left column) and $t = 10^5$ (right column). The first and 
         third rows show the particle monolayer of bijels stabilized with oblate and prolate ellipsoids respectively with no applied field. The second and fourth 
         rows show the particle monolayer of bijels stabilized with oblate and prolate ellipsoids respectively with an applied field of $\bar{B}_z = 1$. Particle 
         reorientation to the direction of the field can be seen, resulting in coarsening of the domains.} 
\label{fig:particle_viz-field_on} 
\end{figure}

In Figure \ref{fig:particle_viz-field_on}, we see the initial configuration of particles at the monolayer show that the orientations are random and that the 
particles tend to lie flat on the interface. However, upon application of the magnetic field the particles orient to the field direction. With no applied field,
the orientations and positions of the particles vary slightly for both ellipsoidal particle morphologies. This effect is due steric effects between particles,
as identified in Gunther et al. \cite{gunther_timescales_2014} Bijels stabilized by prolate ellipsoids, adopt orientational order
to the magnetic field and qualitatively different particle arrangements on the interface. These indicate that the interfacial capillary forces or steric interactions 
with other particles differ upon application of the magnetic field. The particles would prefer to lie flat on the interface to minimize the interfacial energy of 
adsorption onto the interface. Changes in the local arrangement arise from the unjamming and rejamming process upon application of the magnetic field. These differences 
in the orientational and interfacial ordering can be characterized to understand how each respective interaction force changes as a function of time.

>>>>

\section{Results}\label{sec:results_p2}
\subsection{Hysteresis curve}
\label{section:hysteresis_curve}

Bijels have proposed applications as soft-matter-based material templates, drug delivery vectors, and separation systems. In all cases, material performance is 
tightly coupled to the internal microstructure of the bijel template. Applications that benefit from adaptive permeability or structural tunability would be 
enhanced by the ability to reversibly modulate microstructure in situ. This section explores whether magnetically responsive bijels can undergo reversible 
structural changes upon repeated application and removal of an external magnetic field.

We begin by analyzing bijels stabilized by prolate ellipsoidal particles at a particle volume fraction of \(\phi_p = 0.1\). This system is selected due to the 
similar qualitative response mechanisms observed for both prolate and oblate geometries in preliminary simulations (Aim 1), with prolate particles chosen here 
for their greater experimental accessibility. The bijel is initially formed without a magnetic field. After equilibration, a static magnetic field is applied 
in the \(z\)-direction and incrementally increased: \(\bar{B}_z = 0 \rightarrow 0.2 \rightarrow 0.5 \rightarrow 1\). This is followed by a symmetric 
decrement: \(\bar{B}_z = 1 \rightarrow 0.5 \rightarrow 0.2 \rightarrow 0\). At each step, the system is allowed to evolve to a new equilibrium, defined as
the plateauing of the length scale obtained from the domain size, \(L\), computed from the first moment of the spherically averaged structure factor.
\cite{kendon_3d_1999,kendon_inertial_2001} The structure factor is calculated from the density fields as,

%
\begin{equation}
S(\vec{k},t) = \tilde{\phi}(\vec{k},t)\tilde{\phi}(-\vec{k},t) .
\end{equation}
%
where \(\tilde{\phi}(\vec{k},t)\) is the Fourier
transform of the order parameter fluctuations
\(\phi(\vec{x},t)-\left\langle\phi\right\rangle\). Spherical averaging is 
done by averaging the values of the structure factor where $k = |\vec{k}|$ is in
the same bin, defined as

%
\begin{equation}
S(k,t) = \frac{1}{n_k} \sum_{k-\frac{\Delta}{2}<k<k+\frac{\Delta}{2}} \tilde{\phi}(\vec{k},t)\tilde{\phi}(-\vec{k},t) ,
\end{equation}

The lengthscale is then defined as

\begin{equation}
    L = \frac{2 \pi}{\frac{ \sum_k S(k, t)k }{\sum_k S(k, t)}}
\end{equation}

The hysteresis behavior is shown in Figure~\ref{fig:hysteresis_curve}, which plots the normalized domain size \(L/R\) as a function of the applied magnetic 
field at the end of the equilibration step.

@@

The hysteresis behavior is shown in Figure~\ref{fig:hysteresis_curve} demonstrates that as the field is increased, domain size increases, indicating coarsening 
of the fluid domains in the bijel. Karthikeyan and Schiller showed that application of magnetic fields cause reordering of particles at interfaces towards the
direction of the applied field, making the particle monolayer have nematic ordering. In the case of already formed bijels,
this reorientation likely reduces the interface area stabilized, facilitating local unjamming and domain growth. In contrast, when the field strength is reduced, 
the domain size does not revert to its original value, suggesting that the microstructure remains kinetically arrested in the coarsened configuration. This 
irreversibility confirms a hysteretic response of the bijel structure to magnetic field cycling.

A similar phenomenon has been reported by Cui et al.~\cite{cui_stabilizing_2013} in studies of particle-stabilized emulsions subjected to electric fields, 
where spherical particles unjammed and subsequently rejammed in a new configuration, resulting in anisotropic droplet deformation. Our results extend this 
concept to jammed, bicontinuous systems stabilized by anisotropic particles. Notably, Figure~\ref{fig:hysteresis_curve} also shows that the degree of 
structural change is field-strength dependent, indicating that stronger fields drive more extensive particle reconfiguration. In the following sections, 
we investigate how the magnitude of the applied field and the initial degree of particle ordering influence the extent and reversibility of microstructural 
evolution in magnetically responsive bijels.

\subsection{Switching a field onto a bijel with disordered particles}
\subsubsection{Field strength dependence on domain size}
\label{section:field-strength-dependence-on-domain-size}

The results presented in Figure~\ref{fig:hysteresis_curve} demonstrate that changes in bijel domain size are dependent on the strength of the applied magnetic 
field. For applications where bijels serve as porous material templates, microstructural characteristics directly govern performance metrics such as permeability, 
selectivity, and mechanical stability. Furthermore, in many proposed stimuli-responsive applications, understanding not only the extent but also the mechanism
of structural response is critical for functionality and control. In this section, we investigate the dynamic structural response of bijels stabilized by randomly 
oriented ellipsoidal particles with a volume fraction of \(\phi_p = 0.1\), initially formed without a magnetic field. Magnetic fields of varying strengths 
(\(\bar{B} = 0, 0.2, 0.5, 1\)) are applied to these pre-formed structures to assess how the microstructure evolves under external stimuli. We begin by visualizing 
the microstructure of bijels stabilized by prolate particles before and after field application.

\begin{figure} 
\centering 
\includegraphics[scale=0.4]{../figures/results/paper2/microstructure_viz-field_on.png} 
\caption{Visualizations of bijels stabilized by prolate particles simulated under no fields at $t = 0$ (left columns) and $t = 10^5$ (right columns). The top 
         row detail the microstructure evolution for bijels with no applied field while the bottom rows show bijels stabilized by oblate and prolate particles 
         respectively with a field strength of $\bar{B} = 1$ applied.}
\label{fig:microstructure_viz-field_on} 
\end{figure}

Figure~\ref{fig:microstructure_viz-field_on} shows that at zero field, the fluid domains exhibit an isotropic, co-continuous morphology with particles randomly 
distributed and oriented at the fluid interface. Upon applying a magnetic field (\(\bar{B}_z = 1\)), the particles undergo reorientation along the field direction, 
which visibly alters the interfacial configuration and the domain morphology. 
To quantitatively characterize this response, we compute the average domain size as a function of time and magnetic field strength. Additionally, to assess 
how particle alignment contributes to the structural transition, we overlay the evolution of the nematic order parameter with the domain size. This correlation 
provides insight into the mechanistic role of particle ordering in driving microstructural transformation under applied magnetic fields \cite{veerman_phase_1992}.
The nematic order parameter is calculated from the orientations of the particles as

\begin{equation}
\tens{Q} = \frac{1}{n_p} \sum_i \left( \frac{3}{2} \hat{\vec{o}}_i \otimes \hat{\vec{o}}_i -\frac{1}{2}\mathsf{1} \right) ,
\end{equation}

where \(\hat{\vec{o}}_i\) is the orientation of the \(i\)-th particles and \(n_p\) is the number of particles in the system. The largest eigenvalue of 
\(\tens{Q}\) yields the nematic order parameter

\begin{equation}
S = \left\langle \frac{3}{2}\cos^2 (\theta) - \frac{1}{2} \right\rangle ,
\end{equation}

where \(\theta=\arccos(\hat{\vec{o}}_i\cdot\hat{\vec{n}})\) is the angle between the particle orientation and the nematic director \(\hat{\vec{n}}\). We plot $S$
and $L$ over time in Figure \ref{fig:domain_size-field_on}

\begin{figure} 
\centering 
\includegraphics[scale=0.5]{../figures/results/paper2/domain_size-field_on.png} 
\caption{Plots of the time evolution of the normalized domain size \(L/R\) (markers) and nematic order parameter 
         \(S\) (lines) for bijels stabilized by oblate (left, \(\alpha = 0.5\)) and prolate (right, \(\alpha = 2.0\)) ellipsoidal particles under varying 
         magnetic field strengths. Here, \(R = 7.9\) is the volume-equivalent sphere radius.} 
\label{fig:domain_size-field_on} 
\end{figure}

The results show that applying a magnetic field to bijels 
initially formed without one induces domain coarsening, with the final domain size increasing by up to 5.2\% for oblate and 2.5\% for 
prolate particle-stabilized systems.
In both oblate and prolate systems, domain coarsening follows a characteristic three-stage evolution: an initial slow growth phase, a period of rapid 
coarsening, and eventual plateauing as the structure reaches a new jammed configuration. Even in the absence of a magnetic field (\(\bar{B}_z = 0\)), 
domain growth is observed driven by steric rearrangement of interfacial particles, consistent 
with findings by Günther et al.~\cite{gunther_timescales_2014}. Notably, the initiation of domain coarsening under applied magnetic fields is more rapid 
for bijels stabilized by oblate particles compared to prolate systems. However, the rejamming process takes longer.

In the absence of magnetic fields, a slight increase in $S(t)$ is observed, consistent with the sterically driven reorientation of particles at the interface.
\cite{gunther_timescales_2014} Upon application of a magnetic field, the curves for \(S(t)\) increase before plateauing, indicating the parameter has reached
its saturation value. Notably at stronger field strengths, $S(t)$ precedes the increase in domain size for both particle shapes analyzed. This time lag suggests 
that nematic ordering is a precursor, but not the sole driver, of domain coarsening. Overall, these results confirm that domain coarsening in magnetically 
responsive bijels occurs as a result of applied magnetic fields, with the degree of change observed dependent on the strength of the applied field.

\subsubsection{Particle reorientation at the interface}

Building on the observed time lag between the onset of particle alignment and domain coarsening, we next investigate the underlying dynamics of particles at the 
interface. While nematic ordering appears to precede structural changes, the delayed increase in domain size at higher magnetic field strengths suggests that 
additional factors influence the system's response. In this context, capillary interactions between particles and the fluid interface, as well as steric hindrance 
from neighboring particles, may act to resist or slow particle rearrangement, thereby affecting the timescale of domain evolution. To better understand these 
phenomena, we focus on the dynamic behavior of the interfacial particle monolayer. As a first step, we visualize the initial and final configurations of bijels 
stabilized by prolate ellipsoids before and after application of a magnetic field, providing qualitative insight into particle displacement and reorganization 
at the interface.

\begin{figure} 
\centering 
\includegraphics[scale=0.2]{../figures/results/paper2/particle_viz-field_on.png} 
\caption{Visualizations of bijels stabilized by prolate ellipsoidal particles at the initial and final timesteps, 
         both in the absence of a magnetic field (top row) and with a magnetic field applied (\(\bar{B}_z = 1\), bottom row). Particle reorientation to the
         applied field can be observed.} 
\label{fig:particle_viz-field_on} 
\end{figure}

The snapshow shows that in the absence of a magnetic field, minor variations in particle orientation and position occur due to steric interactions, as 
previously observed by Günther et al.~\cite{gunther_timescales_2014}. However, the application of a magnetic field induces clear orientational ordering of the 
particle monolayer, with particles aligning along the field direction and reorganizing into distinct interfacial arrangements. These changes suggest that both 
capillary forces and steric hindrance are modified upon field application. While particles naturally prefer to lie flat at the interface to minimize interfacial 
adsorption energy, the applied field competes with this tendency, reorienting the particles and disrupting their initial packing.

To quantify the orientation of the particles relative to the interface, we analyzed the angle between each particles symmetry axis and 
the local interface normal. We generated a mesh of the interface using the marching cubes algorithm and, for each particle, identified the nearest mesh vertex 
(excluding those beyond the length of the particle's long axis). We then calculated the angle \(\psi\) between the particle orientation vector 
\(\hat{\vec{o}}_i\) and the local surface normal. We plot the time evolution of $\psi$ in Figure \ref{fig:interface_angle-field_on}.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.5]{../figures/results/paper2/psi-field_on.png} 
    \caption{Plots of the evolution of the average interface angle \(\langle \psi \rangle\) over time for bijels stabilized 
             by oblate (left) and prolate (right) ellipsoidal particles under varying magnetic field strengths. The response of \(\langle \psi \rangle\) 
             differs notably between the two particle morphologies.} 
    \label{fig:interface_angle-field_on} 
\end{figure}

For bijels stabilized by oblate particles, \(\langle \psi \rangle\) initially increases from approximately \(19^\circ\), indicating that the particles, which begin 
in an approximately flat configuration at the interface, begin to tilt out of the plane in response to the applied field. After a brief decrease or plateau, the angle 
resumes increasing before stabilizing. This behavior suggests that particles first reorient under magnetic torque but are temporarily resisted by interfacial tension 
and capillary constraints. Once the interface begins to deform in response to particle tilt, the system eventually jams in a new configuration, locking the particles 
at an elevated angle. Both the rate of initial tilt and the duration of the plateau phase increase with field strength, and the final value of 
\(\langle \psi \rangle\) is positively correlated with the magnitude of the applied field.

For bijels stabilized by prolate particles, a similar but more subtle trend is observed. The initial value of \(\langle \psi \rangle\) starts around \(10^\circ\), 
gradually increases, and then exhibits a small plateau followed by a slight decline. While the rate and magnitude of the initial increase are field-dependent, 
the final values converge more closely across different field strengths, suggesting that the field sensitivity of interface tilt is weaker for prolate particles. 
The overall changes in \(\langle \psi \rangle\) are also smaller in magnitude compared to those in oblate systems, indicating that prolate particles experience 
less pronounced reorientation at the interface.

While the interface angle \(\langle \psi \rangle\) provides insight into the out-of-plane orientation of particles in response to magnetic field application, 
it does not capture how particles reorganize at the interface, observed in Figure \ref{fig:microstructure_viz-field_on} Since the evolution of microstructure is 
influenced not only by tilt but also by the steric effects of interfacial particles, we also assess the Steinhardt bond orientational parameter. 
We compute the six fold bond orienational parameter $q_6$ which measures the degree of hexagonal arrangement of particles at the interface.
\cite{steinhardt_bond-orientational_1983, lechner_accurate_2008, mickel_shortcomings_2013} These parameters characterize the local structural ordering by projecting 
the orientations of neighboring particles onto spherical harmonics of degree \(l\), and are defined as

\begin{equation}
q_l(i) = \left( \frac{4\pi}{2l+1} \sum_{m=-l}^{l} \left| \frac{1}{N(i)} \sum_{j=1}^{n(i)} Y_{lm}(\vec{r}_{ij}) \right|^2 \right)^{\frac{1}{2}} ,
\end{equation} 
%\TODO{update formula to the one actually used}

Here, \(n(i)\) denotes the number of nearest neighbors of particle \(i\), \(\vec{r}_{ij}\) is the vector connecting the centers of particles \(i\) and \(j\), and 
\(m\) is an integer ranging from \(-l\) to \(l\) that indexes the spherical harmonic components \(Y_{lm}\). Neighbor lists are typically defined using a fixed cutoff 
radius around each particle, identifying nearby particles as neighbors. We calculate the python package Freud to calculate the bond order parameters and use a Voronoi 
cell to determine the particle neighborhood rather than the standard distance criterion \cite{ramasubramani_freud_2020,mickel_shortcomings_2013}. The Voronoi-based 
method avoids the need for a predefined cutoff and yields more robust neighbor definitions, particularly in disordered systems. Bond orientational order parameters 
have been widely employed to study structural transitions in colloidal and ceramic systems, including crystallization, nucleation, and glass formation 
\cite{vagberg_glassiness_2011, besseling_three-dimensional_2007, schall_structural_2007, ozawa_jamming_2012} We plot the time evolution of $\psi$ in Figure 
\ref{fig:Q6-field_on}.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.5]{../figures/results/paper2/Q6-field_on.png} 
    \caption{Plots of the time evolution of the six-fold Steinhardt bond order parameter \(\langle Q_6 \rangle\) for bijels 
             stabilized by oblate (left) and prolate (right) ellipsoidal particles under varying magnetic field strengths. The dynamics of interfacial ordering differ 
             significantly between particle morphologies with \(\langle Q_6 \rangle\) decreasing over time for oblate particles and increasing for prolate particles.} 
    \label{fig:interface_angle-field_on} 
\end{figure}

From Figure \ref{fig:interface_angle-field_on} the time evolution of \(\langle Q_6 \rangle\) can be qualitatively divided into three regimes;
an initial transition period, a plateau, and a final reordering or jamming phase.
For oblate particles, application of the magnetic field initially causes a sharp decrease in \(\langle Q_6 \rangle\), reflecting a disruption of the 
pre-existing interfacial order as particles begin to tilt and rearrange. This is followed by a plateau, during which ordering is temporarily suppressed 
as particles continue reorienting. Finally, \(\langle Q_6 \rangle\) begins to increase gradually, indicating partial recovery of local order as the system 
evolves toward a new jammed state. The rate of change of \(\langle Q_6 \rangle\) significantly decreases toward the end of the simulation, suggesting the onset 
of kinetic arrest. The final value of \(\langle Q_6 \rangle\) is dependent on the applied magnetic field strength, with higher field strengths 
resulting in lower final ordering, potentially due to increased variance in the interfacial angle, reducing packing efficiency.

In contrast, for bijels stabilized by prolate particles, \(\langle Q_6 \rangle\) increases monotonically with time after the field is applied. The degree 
of ordering and the rate of increase both scale positively with magnetic field strength. The early stage is characterized by a modest rise in \(\langle Q_6 \rangle\), 
followed by a more rapid growth and eventual plateau, indicating progressive in-plane reordering of the particle monolayer as alignment along the field direction 
becomes dominant. The timing and extent of these regimes are also field-dependent with stronger fields result in earlier onset of ordering and higher final values 
of \(\langle Q_6 \rangle\). Notably, for all applied fields, the final ordering significantly exceeds that of the zero-field case. These results highlight that the 
response of interfacial particle order to magnetic fields is morphology-specific as oblate particles undergo a disruption and partial recovery of order, 
while prolate particles experience continuous enhancement of local ordering.

\subsubsection{Microstructural effects of particle reordering}

Karthikeyan and Schiller identified that the reorientation of particles at the interface due to the application of magnetic fields caused domain size anisotropy
driven by the directional cessation of domain coarsening. \cite{karthikeyan_formation_2024} In this study, we have established that the mechanism of domain size change
arises from magnetic field driven rotations of particles at the interface, causing unjamming of the particle monolayer before rejamming and locking in the new
microstructure. Due to the differences in mechanism, we turn to investigate the anisotropic microstructure of the bijel at the final timestep.

We probe the domain anisotropy using the directional second moment of the structure factor, computed along each Cartesian axis from the full structure factor shown
below. \cite{jansen_bijels_2011, gunther_timescales_2014}. 
    %
    \begin{equation}
    L_\beta(t)=2\pi\sqrt{\frac{\sum_{\vec{k}}S(\vec{k},t)}{\sum_{\vec{k}}k_\beta^2 S(\vec{k},t)}} .
    \end{equation}
    %
    This allows us to compute a domain size parallel
    (\(L_\parallel=L_z\)) and perpendicular (\(L_\perp=(L_x+L_y)/2\)) to the
    direction of \(\vec{B}\) separately. The average domain size \(L_d(t)\)
    is computed as \(L_d(t)=\sum_\beta L_\beta(t) / 3\).
    %
    \begin{equation}
    L_d(t)=2\pi\sqrt{\frac{\sum_{\vec{k}}S(\vec{k},t)}{\sum_{\vec{k}}\sum_\beta k_\beta^2 S(\vec{k},t)}} .
    \end{equation}
    %
Previous studies have also shown that a domain size increase is usually accompanied by a decrease in
tortuosity. \cite{karthikeyan_formation_2024} To determine whether these structural 
changes impact macroscopic transport properties, we also analyzed the tortuosity of the bijels. We utilize the package \texttt{taufactor} that implements
a periodic boundary condition solver and calculates the diffusive tortuosity of the bijel structures. Porous structures were 
extracted by binarizing the final order parameter field using a zero threshold, isolating the largest connected (percolating) domain. We plot the results of
the directional domain size and tortuosity in Figure \ref{fig:domain_size_aniso-field_on}

@@

Figure~\ref{fig:domain_size_aniso-field_on} shows that the application of a magnetic field induces clear domain size anisotropy in bijels stabilized by both oblate 
and prolate particles. Even at zero field, a small degree of anisotropy is observed, which increases substantially with rising field strength. For oblate particles, 
the domain sizes perpendicular (\(L_\perp\)) and parallel (\(L_\parallel\)) to the applied field increase by approximately 73\% and 7\%, respectively. In contrast, 
for prolate particles, \(L_\perp\) increases by 10\%, while \(L_\parallel\) increases by 44\%. These results indicate that the dominant direction of coarsening 
depends on particle morphology: oblate particles promote greater coarsening perpendicular to the field, while prolate particles favor coarsening parallel to the 
field. This trend is consistent with previous findings and reflects how particles reorient in response to the applied magnetic field. As particles unjam and rejam 
at the interface, they adopt morphology- and field-dependent orientations that result in directionally biased surface coverage, manifesting as anisotropic domain 
growth.

The right panel of Figure~\ref{fig:domain_size_aniso-field_on} presents the corresponding changes in directional tortuosity. Initially, the tortuosity values 
for both directions are close to \(\tau \approx 1.5\), aligning with previous simulations of gyroidal and co-continuous structures using Lattice Boltzmann methods 
\cite{luo_macroscopic_2020}. Upon applying the magnetic field, \(\tau_\perp\) decreases while \(\tau_\parallel\) increases. This inverse relationship between 
tortuosity and domain size is consistent with prior findings \cite{karthikeyan_formation_2024}, where larger domains exhibited 
reduced tortuosity. The observed anisotropic response suggests that magnetic fields not only influence interfacial particle orientation but also lead to 
direction-dependent transport pathways in the bijel. Moreover, the extent and direction of anisotropy are strongly influenced by the particle morphology, 
particularly the axis of symmetry, which dictates the dominant direction of microstructural reorganization under field application.

Thus far, three key particle-scale phenomena have been characterized in response to the application of a magnetic field. The first is
particle alignment to the field direction, where both oblate and prolate ellipsoids reorient their major axes along the magnetic field vector. The second effect is
changes in particle tilt relative to the fluid interface, reflected in variations of the interface angle \(\langle \psi \rangle\), indicating out of interface 
reorientation driven by magnetic torque. Third, reorganization of the local particle monolayer, as captured by the evolution of the Steinhardt bond order parameter 
\(\langle Q_6 \rangle\), which reflects field-dependent disruptions and recoveries in particle ordering at the interface. Together, these processes 
drive the unjamming of the particle monolayer, causing coarsening of the fluid domains before the interface rejams in place with anisotropy dictated by the orientation
of particles to the magnetic field.

\subsection{Increasing the applied field onto a bijel made with a magnetic field}
\subsubsection{Field strength dependence on domain size}
\label{section:increasing-the-applied-field}

The hysteresis curve presented earlier in Figure~\ref{fig:hysteresis_curve} revealed that microstructural changes—quantified by domain size—diminish 
with each successive increase in magnetic field strength. This saturation behavior suggests a direct link between the extent of domain coarsening 
and the evolving order within the interfacial particle monolayer. As demonstrated in the previous section, once the bijel is jammed, further domain 
growth is constrained by local rearrangements of the particles at the interface, which are themselves influenced by the degree of particle alignment 
and packing. To investigate this relationship further, we now examine how pre-existing particle order, established during bijel formation, affects 
the system's responsiveness to a subsequent applied magnetic field.
